---
title: "CS 586: Milestone 4"
subtitle: "Cleaning"
author:
  - William Baker-Robinsons
  - Kevin Ng
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
---

## Data cleaning associated with the spreadsheet containing information regarding students' availabilities

### Includes creation of the student table, availability table, availability_rel table, and availability_comments table

```{r, warning = F, code = readLines("scheduling.R")}

source("./scheduling.r")

```

## Data cleaning associated with the spreadsheet containing information regarding students' project interests

### Includes creation of the project table and the project_rel table

```{r, warning = F, code = readLines("projects.R")}

source("./projects.r")

```

## Data cleaning associated with the spreadsheet containing students' grades

### Includes creation of the course table and course_rel table

```{r, warning = F, code = readLines("grades.R")}

source("./grades.r")

```

## Updating the tables based on the most recent .csv file, taking into account redundant submissions 

```{r, warning = F, code = readLines("updated.R")}

source("./updated.r")

```