---
title: "Capstone Survey Analysis"
author: "William Baker-Robinson and Kevin Ng"
date: "3/7/2020"
output:  
 html_document:
    theme: lumen
    toc: yes
    toc_float:
      collapsed: yes
      smooth_scroll: no
params:
  availSurv: NA
  projSurv: NA
  eligible: NA
  projInfo: NA
  host: NA
  dbName: NA
  username: NA
  password: NA
  rendered_by_shiny: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r param vars for python}
host <- params$host
dbName <- params$dbName
username <- params$username
password <- params$password
renderVal <- 3
```

```{python connect to database}
import psycopg2
import pandas
import io

connected = True
dbData = False
student = 0

try:
    conn = psycopg2.connect(
      dbname = r.dbName,
      user = r.username,
      host = r.host,
      password = r.password
    )
except:
    print("ERROR: Unable to connect to database")
    connected = False
if(connected):
    conn.autocommit = True
    cur = conn.cursor()
    cur.execute("SELECT 1 FROM information_schema.tables WHERE table_name = 'student';")
    if(cur.fetchone() is None):
      ddl_run = 0
    else:
      ddl_run = 1
    data_exists = []
    has_data = 0
    # fix with None
    if(ddl_run):
      cur.execute("SELECT 1 FROM project;")
      if(cur.fetchone() is None):
        has_data = 0
      else:
        has_data = 1
      data_exists.append(has_data)
      cur.execute("SELECT 1 FROM grade;")
      if(cur.fetchone() is None):
        has_data = 0
      else:
        has_data = 1
      data_exists.append(has_data)
    else:
      data_exists.append(0)
      data_exists.append(0)
```

```{r check user uploads and clean}
UpdateProgress(renderVal, params$rendered_by_shiny)
#check to see if the user uploaded any docs
userData <- FALSE
availSurvBool <- FALSE
projSurvBool <- FALSE
eligibleBool <- FALSE

if(py$connected == TRUE){
  print("Connected to the DB this is where I am going to clean")
  if(py$ddl_run == 0 & !is.null(params$availSurv))
  {
    cleanAvailVec <- scheduling(params$availSurv)
    availSurvBool <- TRUE
  }
  if(!is.null(params$availSurv) | py$ddl_run == 1)
  {
    if(py$data_exists[1] == 0 & !is.null(params$projSurv) & !is.null(params$projInfo))
    {
      cleanProjVec <- projects(params$projSurv, params$projInfo) 
      projSurvBool <- TRUE
    }
    if(py$data_exists[2] == 0 & !is.null(params$eligible))
    {
      cleanGradeVec <- grading(params$eligible)
      eligibleBool <- TRUE
    }
  }
}
UpdateProgress(renderVal, params$rendered_by_shiny)
```

```{r clean data frames into R objects}
if(availSurvBool == TRUE)
{
  student_df <- cleanAvailVec[[1]]
  avail_df <- cleanAvailVec[[2]]
  avail_rel_df <- cleanAvailVec[[3]]
  avail_comm_df <- cleanAvailVec[[4]]
}
if(projSurvBool == TRUE)
{
  interst_df <- cleanProjVec[[1]]
  skill_df <- cleanProjVec[[2]]
  work_style_df <- cleanProjVec[[3]]
  project_rel_df <- cleanProjVec[[4]]
  role_df <- cleanProjVec[[5]]
  interested_in_df <- cleanProjVec[[6]]
  skilled_in_df <- cleanProjVec[[7]]
}
if(eligibleBool == TRUE)
{
  grade_df <- cleanGradeVec[[1]]
  grade_rel_df <- cleanGradeVec[[2]]
}
```

```{python function to change pandas df to file like object for Copy to}
def to_file_obj( dataframe ):
  s_buf = io.StringIO()
  dataframe.to_csv(s_buf, index = False, header = False)
  s_buf.seek(0)
  return s_buf;
```

```{python convert float to int in dfs}
if(r.availSurvBool == True)
  student_df = r.student_df.astype({'student_id':'int'})
  avail_df = r.avail_df.astype({'availability_id':'int'})
  avail_rel_df = r.avail_rel_df.astype('int')
  avail_comm_df = r.avail_comm_df.astype({'student_id':'int'})
if(r.projSurvBool == True)
  interest_df = r.interst_df.astype({'student_id':'int'})
  skill_df = r.skill_df({'student_id':'int'})
  work_style_df = r.work_style_df({'student_id':'int'})
  project_rel_df = r.project_rel_df.astype({'project_id':'int', 'student_id':'int', 'interest':'int', 'confidence':'int'})
  role_df = r.role_df.astype({'student_id':'int', 'role_id':'int'})
  interested_in_df = r.interested_in_df.astype({'student_id':'int'})
  skilled_in_df = r.skilled_in_df.astype({'student_id':'int'})
if(r.eligibleBool == True)
  grade_df = r.grade_df({'student_id':'int'})
  grade_rel_df = r.grade_rel_df.astype({'student_id':'int', 'grade_id':'int'})
```

```{python update or upload}
if (connected == True and r.availSurvBool == True and ddl_run == False):
  print("Connected, user supplied data, time to run DDL")
  commands = (
  """  CREATE TABLE student (
  student_id integer PRIMARY KEY,
  name varchar(50) NOT NULL,
  email varchar(50) UNIQUE NOT NULL
  );
  """,
  """ CREATE TABLE availability(
  availability_id integer PRIMARY KEY,
  day_of_week integer NOT NULL,
  mtg_time integer NOT NULL,
  description integer NOT NULL,
  CHECK (description == 1 OR description == 0),
  CHECK (day_of_week >= 0 AND day_of_week <= 6),
  CHECK (mtg_time >= 0 AND mtg_time <= 17)
  );
  """,
  """ CREATE TABLE availability_rel(
  student_id integer REFERENCES student(student_id),
  availability_id integer REFERENCES availability(availability_id),
  PRIMARY KEY(student_id, availability_id)
  );
  """,
  """ CREATE TABLE availability_comments(
  student_id integer REFERENCES student(student_id),
  avail_comment text,
  unavailability text,
  other text,
  PRIMARY KEY(student_id)
  );
  """,
  """ CREATE TABLE project(
  project_id integer PRIMARY KEY,
  organization varchar(50),
  contact_name text,
  contact_email text,
  description text
  );
  """,
  """ CREATE TABLE project_rel(
  project_id integer REFERENCES project(project_id),
  student_id integer REFERENCES student(student_id),
  interest integer NOT NULL,
  confidence integer NOT NULL,
  comment text,
  CHECK (interest >= 0 AND interest <= 10),
  CHECK (confidence >= 0 AND confidence <= 5)
  );
  """,
  """ CREATE TABLE grade_rel(
  student_id integer REFERENCES student(student_id),
  grade_id integer REFERENCES grade(grade_id)
  PRIMARY KEY(student_id, grade_id)
  );
  """,
  """ CREATE TABLE grade(
  grade_id integer PRIMARY KEY,
  course integer NOT NULL,
  grade_value integer NOT NULL,
  CHECK (course >= 0 AND course <= 11),
  CHECK (grade_value >= 0 AND grade_value <= 21)
  );
  """,
  """ CREATE TABLE interested_in(
  student_id integer REFERENCES student(student_id),
  interested_id integer NOT NULL,
  comment text,
  CHECK (interested_id >= 0 AND interested_id <= 7)
  );
  """,
  """ CREATE TABLE skilled_in(
  student_id integer REFERENCES student(student_id),
  familiar_id integer NOT NULL,
  comment text,
  CHECK (familiar_id >= 0 AND familiar_id <= 7)
  );
  """,
  """ CREATE TABLE roll(
  student_id integer REFERENCES student(student_id),
  role_id integer NOT NULL,
  comment text,
  CHECK (role_id >= 0 AND role_id <= 4)
  );
  """,
  """ CREATE TABLE work_info(
  student_id integer REFERENCES student(student_id),
  strength text NOT NULL,
  weakness text NOT NULL,
  work integer NOT NULL,
  know_about text,
  other text 
  );
  """,
  """ CREATE TABLE interest(
  student_id integer REFERENCES student(student_id),
  interest_most text NOT NULL,
  interest_least text NOT NULL,
  cs_enjoy text NOT NULL
  );
  """,
  """ CREATE TABLE skill(
  student_id integer REFERENCES student(student_id),
  know_most text NOT NULL,
  know_least text NOT NULL,
  languages text NOT NULL
  );
  """)
  try:
    for command in commands:
      cur.execute(command)
  except (Exception, psycopg2.DatabaseError) as error:
    print(error)
  print("DDL run, now time to upload data!")
  # upload availability survey
  cur.copy_from(to_file_obj(student_df), 'student', sep = ',', null = '')
  cur.copy_from(to_file_obj(avail_df), 'availability', sep = ',', null = '')
  cur.copy_from(to_file_obj(avail_rel_df), 'availability_rel', sep = ',', null = '')
  cur.copy_from(to_file_obj(avail_comm_df), 'availability_comments', sep = ',', null = '')
if (connected == True and ddl_run == True):
  if(r.projSurvBool == True)
    print("project data will be uploaded here")
    cur.copy_from(to_file_obj(student_df), 'student', sep = ',', null = '')
    cur.copy_from(to_file_obj(avail_df), 'availability', sep = ',', null = '')
    cur.copy_from(to_file_obj(avail_rel_df), 'availability_rel', sep = ',', null = '')
    cur.copy_from(to_file_obj(avail_comm_df), 'availability_comments', sep = ',', null = '')
  if(r.eligibleBool == True):
    print("grade data will be uploaded here")
    cur.copy_from(to_file_obj(grade_df), 'student', sep = ',', null = '')
    cur.copy_from(to_file_obj(grade_rel_df), 'availability', sep = ',', null = '')
  else:
    print("Connected, and database is setup. No data to upload or database already has data")
elif(connected == True and r.availSurvBool == False and ddl_run == False):
  print("ERROR: Connected but no data exists in the database, and no data provided to add")
else:
  print("Not connected")
# Check if you can query the table for the questions below
```


# 20 Questions     
## Interst Level  
### Count the number of students interested in each project:  
_number of students with interest level greater than 5_  

### Count the number of times a project was selected as a top 3 choice:  

### Are there more students who marked low interest in all projects than students who marked high interest in all projects?  
_How to define low vs. high interest?_  

### Output a list of the 5 students who have the highest interest level for each project:  

### Create a scatterplot of interest level and confidence to complete each project:  

### What skills are students most interested in?  

## Availability  
### When are most students available?  
_Create a heatmap_

### Plot students time availability stratified by 1st choice of project:  

### Plot students time availability stratified by 2nd choice of project:  

### Plot students time availability stratified by 3rd choice of project:  

### What proportion of students perfer times before 10:00 Am?

### What proportion of students are available past 6:00 Pm? 

## Skills
### What are the 5 most common skills?  

### What are the 5 least common skills?  

### What is the most common skill among those who got straight A’s in their prerequisite courses?  
_Do they have to have an A in all pre-reqs? How many people actually did this?_  

### Return a list of the 5 students who are most confident in their ability to complete each project:  

### Create a histogram of the students confidence level in completing each project:  

### Create a barchart of the skills students are skilled in or familiar with:  

## Eligibility
### How many students do not meet eligibility criteria?   

### Where do skills not match with eligibility criteria?  
_May be hard to match skills and classes taken..._  

### What is the average confidence level of students who do not meet eligibility criteria for each project?  


```{python}
if(connected):
  conn.close()
```

