---
title: "Capstone Survey Analysis"
author: "William Baker-Robinson and Kevin Ng"
date: "3/7/2020"
output:  
 html_document:
    theme: lumen
    toc: yes
    toc_float:
      collapsed: yes
      smooth_scroll: no
params:
  availSurv: NA
  projSurv: NA
  eligible: NA
  projInfo: NA
  host: NA
  dbName: NA
  username: NA
  password: NA
  rendered_by_shiny: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r param vars for python}
host <- params$host
dbName <- params$dbName
username <- params$username
password <- params$password
renderVal <- 4
```

```{python connect to database}
import psycopg2
import pandas

connected = True
dbData = False
student = 0

try:
    conn = psycopg2.connect(
      dbname = r.dbName,
      user = r.username,
      host = r.host,
      password = r.password
    )
except:
    print("ERROR: Unable to connect to database")
    connected = False
if(connected):
    cur = conn.cursor()
    cur.execute("SELECT 1 FROM information_schema.tables WHERE table_name = 'student';")
    student = cur.fetchone()
if(student is not None):
    dbData = True
```

```{r check user uploads and clean}
UpdateProgress(renderVal, params$rendered_by_shiny)
#check to see if the user uploaded any docs
userData <- FALSE
availSurvBool <- FALSE
projSurvBool <- FALSE
eligibleBool <- FALSE
projInfoBool <- FALSE
if(py$connected == TRUE){
  print("Connected to the DB this is where I am going to clean")
  if(!is.null(params$availSurv))
  {
    availabilityDfs <- AvailabilityClean(params$availSurv)
    availSurvBool <- TRUE
  }
  if(!is.null(params$projSurv))
  {
    projectDfs <- ProjectsClean(params$projSurv)
    projSurvBool <- TRUE
  }
  if(!is.null(params$eligible))
  {
    eligibleDfs <- GradesClean(params$eligible)
    eligibleBool <- TRUE
  }
  if(!is.null(params$projectInfo))
  {
    # not sure about where the projectInfo doc is cleaned
    projectInfoBool <- TRUE
  }
  if(availSurvBool | projSurvBool | eligibleBool | projInfoBool)
  {
    userData <- TRUE
  }
}
UpdateProgress(renderVal, params$rendered_by_shiny)
```

```{python update or upload}
canQuery = False

if (connected == True and r.userData == True and dbData == False):
  print("Connected, user supplied data, and db is empty")
  # run ddl 
  # upload all...
  canQuery = True
elif (connected == True and r.userData == True and dbData == True):
  #Will create if have time for now nothing done with data just produces report.
    #find where new data starts in student avail
    #assumption is that only the survey 1 & 2 data will be uploaded and updated like this
    #compare new section to prev to check for duplicate stu id
    # keep track of all id which are duplicated
    # update info for duplicated
    # load the rest to DB
  print("For now this program does not supported the database updates")
  canQuery = True
elif (connected == True and r.userData == False and dbData == True):
  print("connected, user did not supply data, and the db has data")
  canQuery = True  
elif(connected == True and r.userData == False and dbData == False):
  print("ERROR: Connected but no data exists in the database, and no data provided to add")
else:
  print("Not connected")
```


# 20 Questions     
## Interst Level  
### Count the number of students interested in each project:  
_# of students with interest level greater than 5_  

### Count the number of times a project was selected as a top 3 choice:  

### Are there more students who marked low interest in all projects than students who marked high interest in all projects?  
_How to define low vs. high interest?_  

### Output a list of the 5 students who have the highest interest level for each project:  

### Create a scatterplot of interest level and confidence to complete each project:  

### What skills are students most interested in?  

## Availability  
### When are most students available?  
_Create a heatmap_

### Plot students time availability stratified by 1st choice of project:  

### Plot students time availability stratified by 2nd choice of project:  

### Plot students time availability stratified by 3rd choice of project:  

### What proportion of students perfer times before 10:00 Am?

### What proportion of students are available past 6:00 Pm? 

## Skills
### What are the 5 most common skills?  

### What are the 5 least common skills?  

### What is the most common skill among those who got straight A’s in their prerequisite courses?  
_Do they have to have an A in all pre-reqs? How many people actually did this?_  

### Return a list of the 5 students who are most confident in their ability to complete each project:  

### Create a histogram of the students confidence level in completing each project:  

### Create a barchart of the skills students are skilled in or familiar with:  

## Eligibility
### How many students do not meet eligibility criteria?   

### Where do skills not match with eligibility criteria?  
_May be hard to match skills and classes taken..._  

### What is the average confidence level of students who do not meet eligibility criteria for each project?  


```{python}
if(connected):
  conn.close()
```

