---
title: "Capstone Survey Analysis"
author: "William Baker-Robinson and Kevin Ng"
date: "3/7/2020"
output:  
 html_document:
    theme: lumen
    toc: yes
    toc_float:
      collapsed: yes
      smooth_scroll: no
params:
  availSurv: NA
  projSurv: NA
  eligible: NA
  projInfo: NA
  host: NA
  dbName: NA
  username: NA
  password: NA
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r param vars for python}
host <- params$host
dbName <- params$dbName
username <- params$username
password <- params$password
```

```{python connect to database}
import psycopg2
import pandas

connected = True
dbData = False

try:
    conn = psycopg2.connect(
      dbname = r.dbName,
      user = r.username,
      host = r.host,
      password = r.password
    )
except:
    print("Unable to connect to database")
    connected = False
if(connected):
    cur = conn.cursor()
# CHECK IF STUDENT HAS DATA IN IT
if(student == True)
  dbData = True
```

```{r check user uploads and clean}
#check to see if the user uploaded any docs
userData <- FALSE
availSurv <- FALSE
projSurv <- FALSE
eligible <- FALSE
projInfo <- FALSE
if(py$connected == TRUE){
  # check avail surv
    # if true clean
    # set to true
  # check proj surv
    # if true clean
    # set to true
  # check eligible
    # if true clean
    # set to true
  # check projInfo
    # if true clean
    # set to true
  userData <- TRUE
}
```

```{python update or upload}
canQuery = False

if (connected == True and r.userData == True and dbData == False):
  # run ddl 
  # upload all...
  canQuery = True
# elif (connected == True and r.userData == True and dbData == True):
  #Will create if have time
    #find where new data starts in student avail
    #assumption is that only the survey 1 & 2 data will be uploaded and updated like this
    #compare new section to prev to check for duplicate stu id
    # keep track of all id which are duplicated
    # update info for duplicated
    # load the rest to DB
  #canQuery = True
elif (connected == True and r.userData == False and dbData == True):
  canQuery = True  
else:
  print("Not connected")
```

# About
This project was created by Will and Kevin for CS 586. It is designed to help with the analysis of surveys relating to the PSU Senior Capstone Class. Hopefully this document, and the database created will be able to be used for easier selection of project teams, and the summary of other capstone related information.
Some notes about our program:  
(@) This program will create a database of the information supplied on whichiver database you provide credentials for. This should work as long as the database is Postgres.  
(@) After the database is created you will get this document as a .html with answers to the 20 questions.  
(@) If there are no files submited at the start the program will atempt to conenct to the database and run the report.  
(@) The file upload format for the availability survey, project interest survey, and projects information is .csv. We expect excel files for the eligibility information document.  

# 20 Questions     
## Interst Level  
### List students based on project of interest:  

### Count the number of students interested in each project:  

### Which project garnered the least amount of interest?  

### Are there more students who marked low interest in all projects than students who marked high interest in all projects?  

### What is the average amount of interest for project 3?  

### How many students listed more than 5 skills or subjects that interest them the most?  

### Is there at least one pair of students where each student has similar degrees of interest in a project and where a subject that one student feels the most knowledgeable about is the same subject that the other student feels the least knowledgeable about? 

### How many students share a favorite programming language with another student who expresses the same confidence level for at least one project?  

## Availability  
### When are most students available?  

### What proportion of students are available past 6:00 Pm?  

### What proportion of students perfer times before 8:00 Am?

### What is the minimum number of time slot availabilities of any student?  

### Which students have preferred times that are not available times?  

## Skills
### What are the most common skills?  

### What are some of the least common skills?  

### Where do skills not match with eligibility criteria?  

### Of the students who expressed at least moderate interest in over half the projects, what percentage of them also expressed at least moderate confidence in their skill?   

### What is the most common skill among those who got straight A’s in their prerequisite courses?  

## Miscellaneous
### Which student left the most additional comments?  

### Is the proportion of solo workers who responded to the prompt regarding anything else to know about them greater than or less than the proportion of non-solo workers who responded?  

```{python}
if(connected):
  conn.close()
```

