---
title: "Capstone Survey Analysis"
author: "William Baker-Robinson and Kevin Ng"
date: "3/7/2020"
output:  
 html_document:
    theme: lumen
    toc: yes
    toc_float:
      collapsed: yes
      smooth_scroll: no
params:
  availSurv: NA
  projSurv: NA
  eligible: NA
  projInfo: NA
  host: NA
  dbName: NA
  username: NA
  password: NA
  rendered_by_shiny: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r param vars for python}
host <- params$host
dbName <- params$dbName
username <- params$username
password <- params$password
renderVal <- 3
```

```{python connect to database}
import psycopg2
import pandas
import io

connected = True
dbData = False
ddl_run = False

try:
    conn = psycopg2.connect(
      dbname = r.dbName,
      user = r.username,
      host = r.host,
      password = r.password
    )
except psycopg2.Error as error:
    print(error)
    pass
    connected = False
if(connected):
  conn.autocommit = True
  cur = conn.cursor()
  cur.execute("SELECT 1 FROM information_schema.tables WHERE table_name = 'student';")
  if(cur.fetchone() is not None):
    ddl_run = True
  data_exists = []
  has_data = 0
  # fix with None
  if(ddl_run == True):
    cur.execute("SELECT 1 FROM project;")
    if(cur.fetchone() is None):
      has_data = 0
    else:
      has_data = 1
    data_exists.append(has_data)
    cur.execute("SELECT 1 FROM grade;")
    if(cur.fetchone() is None):
      has_data = 0
    else:
      has_data = 1
    data_exists.append(has_data)
  else:
    data_exists.append(0)
    data_exists.append(0)
```

```{r check user uploads and clean}
renderVal <- UpdateProgress(renderVal, params$rendered_by_shiny)
#check to see if the user uploaded any docs
availSurvBool <- FALSE
projSurvBool <- FALSE
eligibleBool <- FALSE

if(py$connected == TRUE){
  print("Connected to the DB this is where I am going to clean")
  if(py$ddl_run == FALSE & !is.null(params$availSurv))
  {
    cleanAvailVec <- scheduling(params$availSurv)
    renderVal <- UpdateProgress(renderVal, params$rendered_by_shiny) # update shiny rendering bar
    availSurvBool <- TRUE
  }
  if(!is.null(params$availSurv) | py$ddl_run == TRUE)
  {
    if(py$data_exists[1] == 0 & !is.null(params$projSurv) & !is.null(params$projInfo))
    {
      cleanProjVec <- projects(params$projSurv, params$projInfo) 
      renderVal <- UpdateProgress(renderVal, params$rendered_by_shiny) # update shiny rendering bar
      projSurvBool <- TRUE
    }
    if(py$data_exists[2] == 0 & !is.null(params$eligible))
    {
      cleanGradeVec <- grading(params$eligible)
      renderVal <- UpdateProgress(renderVal, params$rendered_by_shiny) # update shiny rendering bar
      eligibleBool <- TRUE
    }
  }
}
```

```{r clean data frames into R objects}
if(availSurvBool == TRUE)
{
  student_df <- cleanAvailVec[[1]]
  avail_df <- cleanAvailVec[[2]]
  avail_rel_df <- cleanAvailVec[[3]]
  avail_comm_df <- cleanAvailVec[[4]]
}
if(projSurvBool == TRUE)
{
  interst_df <- cleanProjVec[[1]]
  skill_df <- cleanProjVec[[2]]
  work_style_df <- cleanProjVec[[3]]
  project_rel_df <- cleanProjVec[[4]]
  role_df <- cleanProjVec[[5]]
  interested_in_df <- cleanProjVec[[6]]
  skilled_in_df <- cleanProjVec[[7]]
  project_df <- cleanProjVec[[8]]
}
if(eligibleBool == TRUE)
{
  grade_df <- cleanGradeVec[[1]]
  grade_rel_df <- cleanGradeVec[[2]]
}
renderVal <- UpdateProgress(renderVal, params$rendered_by_shiny) # update shiny rendering bar
```

```{python function to change pandas df to file like object for Copy to}
def to_file_obj( dataframe ):
  s_buf = io.StringIO()
  dataframe.to_csv(s_buf, index = False, header = False, na_rep = 'NULL', sep = '\t')
  s_buf.seek(0)
  return s_buf;
```

```{python convert float to int in dfs}
if(r.availSurvBool == True):
  student_df = r.student_df.astype({'student_id':'int'})
  avail_df = r.avail_df.astype({'availability_id':'int'})
  avail_rel_df = r.avail_rel_df.astype('int')
  avail_comm_df = r.avail_comm_df.astype({'student_id':'int'})
if(r.projSurvBool == True):
  interest_df = r.interst_df.astype({'student_id':'int'})
  skill_df = r.skill_df.astype({'student_id':'int'})
  work_style_df = r.work_style_df.astype({'student_id':'int'})
  project_df = r.project_df.astype({'project_id':'int'})
  project_rel_df = r.project_rel_df.astype({'project_id':'int', 'student_id':'int', 'interest':'int', 'confidence':'int'})
  role_df = r.role_df.astype({'student_id':'int', 'role_id':'int'})
  interested_in_df = r.interested_in_df.astype({'student_id':'int'})
  skilled_in_df = r.skilled_in_df.astype({'student_id':'int'})
if(r.eligibleBool == True):
  grade_df = r.grade_df.astype({'grade_id':'int'})
  grade_rel_df = r.grade_rel_df.astype({'student_id':'int', 'grade_id':'int'})
```

```{python update or upload}
if (connected == True and r.availSurvBool == True and ddl_run == False):
  print("Connected, user supplied data, time to run DDL")
  commands = (
  """  CREATE TABLE student (
  student_id integer PRIMARY KEY,
  email varchar(50) NOT NULL,
  name varchar(50) UNIQUE NOT NULL
  );
  """,
  """ CREATE TABLE availability (
  availability_id integer PRIMARY KEY,
  day_of_week varchar(10) NOT NULL,
  description varchar(10) NOT NULL,
  mtg_time varchar(10) NOT NULL
  );
  """,
  """ CREATE TABLE availability_rel (
  student_id integer REFERENCES student(student_id),
  availability_id integer REFERENCES availability(availability_id),
  PRIMARY KEY(student_id, availability_id)
  );
  """,
  """ CREATE TABLE availability_comments (
  student_id integer REFERENCES student(student_id),
  comment text,
  unavailability text,
  other text,
  PRIMARY KEY(student_id)
  );
  """,
  """ CREATE TABLE project (
  project_id integer PRIMARY KEY,
  organization varchar(150),
  contact_name text,
  contact_email text,
  project_name text,
  description text
  );
  """,
  """ CREATE TABLE project_rel (
  project_id integer REFERENCES project(project_id),
  student_id integer REFERENCES student(student_id),
  interest integer NOT NULL,
  confidence integer NOT NULL,
  comment text,
  CHECK (interest >= 0 AND interest <= 10),
  CHECK (confidence >= 0 AND confidence <= 5)
  );
  """,
  """ CREATE TABLE grade (
  grade_id integer PRIMARY KEY,
  course varchar(6) NOT NULL,
  grade_value varchar(4) NOT NULL
  );
  """,
  """ CREATE TABLE grade_rel (
  student_id integer REFERENCES student(student_id),
  grade_id integer REFERENCES grade(grade_id),
  PRIMARY KEY(student_id, grade_id)
  );
  """,
  """ CREATE TABLE interested_in (
  student_id integer REFERENCES student(student_id),
  interested_id integer NOT NULL,
  comment text,
  CHECK (interested_id >= 0 AND interested_id <= 10)
  );
  """,
  """ CREATE TABLE skilled_in (
  student_id integer REFERENCES student(student_id),
  familiar_id integer NOT NULL,
  comment text,
  CHECK (familiar_id >= 0 AND familiar_id <= 10)
  );
  """,
  """ CREATE TABLE role (
  student_id integer REFERENCES student(student_id),
  role_id integer NOT NULL,
  comment text,
  CHECK (role_id >= 0 AND role_id <= 4)
  );
  """,
  """ CREATE TABLE work_style (
  student_id integer REFERENCES student(student_id),
  strength text NOT NULL,
  weakness text NOT NULL,
  work varchar(18) NOT NULL,
  know_about text,
  other text 
  );
  """,
  """ CREATE TABLE interest (
  student_id integer REFERENCES student(student_id),
  s_interest_m text NOT NULL,
  s_interest_l text NOT NULL,
  enjoyed text NOT NULL
  );
  """,
  """ CREATE TABLE skill (
  student_id integer REFERENCES student(student_id),
  s_knowledge_m text NOT NULL,
  s_knowledge_l text NOT NULL,
  languages text NOT NULL
  );
  """)
  try:
      for command in commands:
        cur.execute(command)
  except psycopg2.DatabaseError as error:
    pass
    print(error)
  #upload availability survey    
  try:
    cur.copy_from(to_file_obj(student_df), 'student', sep = '\t', null = 'NULL')
    cur.copy_from(to_file_obj(avail_df), 'availability', sep = '\t', null = 'NULL')
    cur.copy_from(to_file_obj(avail_rel_df), 'availability_rel', sep = '\t', null = 'NULL')
    cur.copy_from(to_file_obj(avail_comm_df), 'availability_comments', sep = '\t', null = 'NULL')
    ddl_run = True
  except psycopg2.DatabaseError as error:
    pass
    print(error)
if (connected == True and ddl_run == True):
  if(r.projSurvBool == True):
    print("project data will be uploaded here")
    try:
      cur.copy_from(to_file_obj(interest_df), 'interest', sep = '\t', null = 'NULL')
      cur.copy_from(to_file_obj(skill_df), 'skill', sep = '\t', null = 'NULL')
      cur.copy_from(to_file_obj(work_style_df), 'work_style', sep = '\t', null = 'NULL')
      cur.copy_from(to_file_obj(project_df), 'project', sep = '\t', null = 'NULL')
      cur.copy_from(to_file_obj(project_rel_df), 'project_rel', sep = '\t', null = 'NULL')
      cur.copy_from(to_file_obj(role_df), 'role', sep = '\t', null = 'NULL')
      cur.copy_from(to_file_obj(interested_in_df), 'interested_in', sep = '\t', null = 'NULL')
      cur.copy_from(to_file_obj(skilled_in_df), 'skilled_in', sep = '\t', null = 'NULL')
    except psycopg2.DatabaseError as error:
      pass
      print(error)
  if(r.eligibleBool == True):
    print("grade data will be uploaded here")
    try:
      cur.copy_from(to_file_obj(grade_df), 'grade', sep = '\t', null = 'NULL')
      cur.copy_from(to_file_obj(grade_rel_df), 'grade_rel', sep = '\t', null = 'NULL')
    except psycopg2.DatabaseError as error:
      pass
      print(error)
  else:
    print("Connected, and database is setup. No data to upload or database already has data")
elif(connected == True and r.availSurvBool == False and ddl_run == False):
  print("ERROR: Connected but no data exists in the database, and no data provided to add")
else:
  print("Not connected")
# Check if you can query the table for the questions below
```

# 20 Questions     
## Interest Level  
### Count the number of students interested in each project:  
_number of students with interest level greater than 5_  
```{python question 1a}
error = False
try:
  sql = "SELECT project_name, COUNT(*) FROM project_rel R JOIN project P ON (R.project_id = P.project_id) WHERE interest > 5 GROUP BY project_name ORDER BY COUNT(*) DESC;"
  data = pandas.read_sql_query(sql, conn)
except:
  print("Error with query")
  pass
  error = True
```

```{r question 1b}
#py$data is now a data frame that can be used in R
if(py$error == FALSE)
{
  kable(py$data, col.names = c("Project Name", "Count"), caption = "Interest of 5 or greater for each project") %>%
    kable_styling(bootstrap_options = c("striped"), full_width = FALSE)
}
renderVal <- UpdateProgress(renderVal, params$rendered_by_shiny)
```

### Count the number of times a project was selected as a top 3 choice:  
```{python question 2a}
qerror = False
try:
  sql = "<your query here>;"
  data = pandas.read_sql_query(sql, conn)
except:
  print("Error with query")
  qerror = True
  pass
```

```{r question 2b}
if(py$qerror == FALSE)
{
  
}
renderVal <- UpdateProgress(renderVal, params$rendered_by_shiny)
```

### Output a list of the 5 students who have the highest interest level for each project:  
```{python question3a}
qerror = False
try:
  command = """CREATE VIEW student_interest AS (
    SELECT S.student_id, S.name, S.email, PR.interest, P.project_id, P.project_name
    FROM student S, project_rel PR, project P
    WHERE S.student_id = PR.student_id AND PR.project_id = P.project_id
  );"""
  cur.execute(command)
except psycopg2.Error as error:
  print(error)
try:
  sql = "SELECT * FROM ((SELECT name, email, interest, project_name FROM student_interest WHERE project_id = 1 ORDER BY interest DESC LIMIT 5) UNION (SELECT name, email, interest, project_name FROM student_interest WHERE project_id = 2 ORDER BY interest DESC LIMIT 5) UNION (SELECT name, email, interest, project_name FROM student_interest WHERE project_id = 3 ORDER BY interest DESC LIMIT 5) UNION (SELECT name, email, interest, project_name FROM student_interest WHERE project_id = 4 ORDER BY interest DESC LIMIT 5) UNION (SELECT name, email, interest, project_name FROM student_interest WHERE project_id = 5 ORDER BY interest DESC LIMIT 5) UNION (SELECT name, email, interest, project_name FROM student_interest WHERE project_id = 6 ORDER BY interest DESC LIMIT 5) UNION (SELECT name, email, interest, project_name FROM student_interest WHERE project_id = 7 ORDER BY interest DESC LIMIT 5) UNION (SELECT name, email, interest, project_name FROM student_interest WHERE project_id = 8 ORDER BY interest DESC LIMIT 5)) AS A ORDER BY A.project_name ASC;"
  data = pandas.read_sql_query(sql, conn)
except:
  print("Error with query")
  pass
  qerror = True
```

```{r question 3b}
if(py$qerror == FALSE)
{
  kable(py$data, caption = "5 students with highest interest for each project") %>%
    kable_styling(bootstrap_options = c("striped"), full_width = FALSE)
}
renderVal <- UpdateProgress(renderVal, params$rendered_by_shiny)
```

### Create a scatterplot of interest level and confidence to complete each project:  
```{python question 4a}
qerror = False
try:
  sql = "SELECT interest, confidence, project_name FROM project_rel R JOIN project P on (R.project_id = P.project_id);"
  data = pandas.read_sql_query(sql, conn)
except:
  print("Error with query")
  qerror = True
  pass
```

```{r question 4b}
if(py$qerror == FALSE)
{
  
}
renderVal <- UpdateProgress(renderVal, params$rendered_by_shiny)
```

### What skills are students most interested in?  
```{python question 5a}
qerror = False
try:
  sql = "<your query here>;"
  data = pandas.read_sql_query(sql, conn)
except:
  print("Error with query")
  qerror = True
  pass
```

```{r question 5b}
if(py$qerror == FALSE)
{
  
}
renderVal <- UpdateProgress(renderVal, params$rendered_by_shiny)
```

## Availability  
### When are most students available?  
_Create a heatmap_
```{python question 6a}
qerror = False
try:
  sql = "<your query here>;"
  data = pandas.read_sql_query(sql, conn)
except:
  print("Error with query")
  qerror = True
  pass
```

```{r question 6b}
if(py$qerror == FALSE)
{
  
}
renderVal <- UpdateProgress(renderVal, params$rendered_by_shiny)
```

### Plot students time availability stratified by 1st choice of project:  
```{python question 7a}
qerror = False
try:
  sql = "<your query here>;"
  data = pandas.read_sql_query(sql, conn)
except:
  print("Error with query")
  qerror = True
  pass
```

```{r question 7b}
if(py$qerror == FALSE)
{
  
}
renderVal <- UpdateProgress(renderVal, params$rendered_by_shiny)
```

### Plot students time availability stratified by 2nd choice of project:  
```{python question 8a}
qerror = False
try:
  sql = "<your query here>;"
  data = pandas.read_sql_query(sql, conn)
except:
  print("Error with query")
  qerror = True
  pass
```

```{r question 8b}
if(py$qerror == FALSE)
{
  
}
renderVal <- UpdateProgress(renderVal, params$rendered_by_shiny)
```

### Plot students time availability stratified by 3rd choice of project:  
```{python question 9a}
qerror = False
try:
  sql = "<your query here>;"
  data = pandas.read_sql_query(sql, conn)
except:
  print("Error with query")
  qerror = True
  pass
```

```{r question 9b}
if(py$qerror == FALSE)
{
  
}
renderVal <- UpdateProgress(renderVal, params$rendered_by_shiny)
```

### What proportion of students prefer times before 10:00 AM?

```{python question 10a}
qerror = False
try:
  sql = "select (select count(*) * 100.0 from student, availability_rel, availability where student.student_id = availability_rel.student_id and availability_rel.availability_id = availability.availability_id and availability.description = 'Preferred' and (availability.mtg_time = '6:00 am' or availability.mtg_time = '7:00 am' or availability.mtg_time = '8:00 am' or availability.mtg_time = '9:00 am'))/count(*) from student, availability_rel, availability where student.student_id = availability_rel.student_id and availability_rel.availability_id = availability.availability_id;"
  data = pandas.read_sql_query(sql, conn)
except:
  print("Error with query")
  qerror = True
  pass
```

```{r question 10b}
if(py$qerror == FALSE)
{
  print(py$data)
}
renderVal <- UpdateProgress(renderVal, params$rendered_by_shiny)
```

### What proportion of students are available past 6:00 Pm? 

```{python question 11a}
try:
  sql = "select (select count(*) * 100.0 from student, availability_rel, availability where student.student_id = availability_rel.student_id and availability_rel.availability_id = availability.availability_id and availability.description = 'Available' and (availability.mtg_time = '7:00 pm' or availability.mtg_time = '8:00 pm' or availability.mtg_time = '9:00 pm' or availability.mtg_time = '10:00 pm' or availability.mtg_time = '11:00 pm'))/count(*) from student, availability_rel, availability where student.student_id = availability_rel.student_id and availability_rel.availability_id = availability.availability_id;"
  data = pandas.read_sql_query(sql, conn)
except:
  print("Error with query")
  pass
```

```{r question 11b}
if(py$qerror == FALSE)
{
  print(py$data)
}
renderVal <- UpdateProgress(renderVal, params$rendered_by_shiny)

```

## Skills

### What are the 5 most common skills? 

```{python question 12a}
try:
  sql = "select familiar_id, count(*) from student, skilled_in where student.student_id = skilled_in.student_id group by skilled_in.familiar_id order by count desc limit 5;"
  data = pandas.read_sql_query(sql, conn)
except:
  print("Error with query")
  pass
```

```{r question 12b}
if(py$qerror == FALSE)
{
  print(py$data)
}
renderVal <- UpdateProgress(renderVal, params$rendered_by_shiny)

```

### What are the 5 least common skills?  
```{python question 13a}
try:
  sql = "select familiar_id, count(*) from student, skilled_in where student.student_id = skilled_in.student_id group by skilled_in.familiar_id order by count limit 5;"
  data = pandas.read_sql_query(sql, conn)
except:
  print("Error with query")
  pass
```

```{r question 13b}
if(py$qerror == FALSE)
{
  print(py$data)
}
renderVal <- UpdateProgress(renderVal, params$rendered_by_shiny)

```

### What is the most common skill among those who got an A in CS350?

```{python question 14a}
try:
  sql = "select familiar_id, count(*) from student, skilled_in, grade_rel, grade where student.student_id = skilled_in.student_id and grade_rel.student_id = student.student_id and grade.grade_id = grade_rel.grade_id and grade.grade_value = 'A' and grade.course = 'CS350' group by skilled_in.familiar_id order by count desc limit 1;"
  data = pandas.read_sql_query(sql, conn)
except:
  print("Error with query")
  pass
```

```{r question 14b}
if(py$qerror == FALSE)
{
  print(py$data)
}
renderVal <- UpdateProgress(renderVal, params$rendered_by_shiny)

```

### Which student is the most confident in their ability to complete project 2?

```{python question 15a}
try:
  sql = "select a.student_id from student a, project_rel b where a.student_id = b.student_id and b.project_id = 2 and b.confidence in (select max(c.confidence) from project_rel c where c.project_id = 2);"
  data = pandas.read_sql_query(sql, conn)
except:
  print("Error with query")
  pass
```

```{r question 15b}
if(py$qerror == FALSE)
{
  print(py$data)
}
renderVal <- UpdateProgress(renderVal, params$rendered_by_shiny)

```

### Which students failed CS161? 
```{python question 16a}
try:
  sql = "select student.student_id from student, grade_rel, grade where student.student_id = grade_rel.student_id and grade_rel.grade_id = grade.grade_id and grade.course = 'CS161' and grade.grade_value = 'F';"
  data = pandas.read_sql_query(sql, conn)
except:
  print("Error with query")
  pass
```

```{r question 16b}

if(py$qerror == FALSE)
{
  print(py$data)
}
renderVal <- UpdateProgress(renderVal, params$rendered_by_shiny)


```

### Create a barchart of the skills students are skilled in or familiar with:  
```{python question 17a}
try:
  sql = "select familiar_id, count(*) from skilled_in group by familiar_id;"
  data = pandas.read_sql_query(sql, conn)
except:
  print("Error with query")
  pass
```

```{r question 17b}

ggplot2::ggplot(py$data, aes(x = familiar_id, y = count,
                             fill = "blue")) +
  geom_col() +
  labs(title = "Distribution of Skill Familiarity",
       x = "Types of Skills",
       y = "Proportion") +
  scale_y_continuous(labels = function(x) paste0(x, '%')) +
  scale_x_continuous(breaks = 0:9, labels = c("FED", "BED", "SS", "WD", "MD", "ES", "UI", "UX", "None", "Other")) + 
  theme(legend.position = "none")

```

## Eligibility
### How many students do not meet eligibility criteria?   
```{python question 18a}
try:
  sql = "select count(*) from student, grade_rel, grade where student.student_id = grade_rel.student_id and grade_rel.grade_id = grade.grade_id and (grade.course = 'CS300' or grade.course = 'CS320' or grade.course = 'CS333' or grade.course = 'CS350') and (grade.grade_value = 'C-' or grade.grade_value = 'D+' or grade.grade_value = 'D' or grade.grade_value = 'D-' or grade.grade_value = 'F');"
  data = pandas.read_sql_query(sql, conn)
except:
  print("Error with query")
  pass
```

```{r question 18b}
if(py$qerror == FALSE)
{
  print(py$data)
}
renderVal <- UpdateProgress(renderVal, params$rendered_by_shiny)

```

### Which students have the same degree of interest in some project as someone who has at least one of the same skills?  

```{python question 19a}
try:
  sql = "select distinct a.student_id from student a, project_rel b, skilled_in z where a.student_id = b.student_id and z.student_id = a.student_id and b.interest = some (select c.interest from project_rel c, student d, skilled_in f where d.student_id = c.student_id and c.project_id = b.project_id and f.student_id = d.student_id and d.student_id <> a.student_id  and f.familiar_id = z.familiar_id);"
  data = pandas.read_sql_query(sql, conn)
except:
  print("Error with query")
  pass
```

```{r question 19b}

if(py$qerror == FALSE)
{
  print(py$data)
}
renderVal <- UpdateProgress(renderVal, params$rendered_by_shiny)

```

### What is the average level of confidence students have with project 6? 
```{python question 20a}
try:
  sql = "select avg(confidence) from project_rel where project_rel.project_id = 6;"
  data = pandas.read_sql_query(sql, conn)
except:
  print("Error with query")
  pass
```

```{r question 20b}

if(py$qerror == FALSE)
{
  print(py$data)
}
renderVal <- UpdateProgress(renderVal, params$rendered_by_shiny)

```


```{python}
# This is for testing
print("""if (connected == True and r.availSurvBool == True and ddl_run == False):
  cur.execute("DROP TABLE availability_rel;")
  cur.execute("DROP TABLE availability_comments;")
  cur.execute("DROP TABLE interest;")
  cur.execute("DROP TABLE skill CASCADE;")
  cur.execute("DROP TABLE work_style;")
  cur.execute("DROP TABLE project_rel;")
  cur.execute("DROP TABLE project;")
  cur.execute("DROP TABLE role;")
  cur.execute("DROP TABLE interested_in;")
  cur.execute("DROP TABLE skilled_in;")
  cur.execute("DROP TABLE grade_rel;")
  cur.execute("DROP TABLE grade;")
  cur.execute("DROP TABLE availability cascade;")
  cur.execute("DROP TABLE student cascade;") """)
if(connected):
  conn.close()
```

