---
title: "Capstone Survey Analysis"
author: "William Baker-Robinson and Kevin Ng"
date: "3/7/2020"
output:  
 html_document:
    theme: lumen
    toc: yes
    toc_float:
      collapsed: yes
      smooth_scroll: no
params:
  availSurv: NA
  projSurv: NA
  eligible: NA
  projInfo: NA
  host: NA
  dbName: NA
  username: NA
  password: NA
  rendered_by_shiny: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r param vars for python}
host <- params$host
dbName <- params$dbName
username <- params$username
password <- params$password
renderVal <- 4
```

```{python connect to database}
import psycopg2
import pandas

connected = True
dbData = False
student = 0

try:
    conn = psycopg2.connect(
      dbname = r.dbName,
      user = r.username,
      host = r.host,
      password = r.password
    )
except:
    print("ERROR: Unable to connect to database")
    connected = False
if(connected):
    cur = conn.cursor()
    cur.execute("SELECT 1 FROM information_schema.tables WHERE table_name = 'student';")
    if(cur.fetchone() is None):
      ddl_run = 0
    else:
      ddl_run = 1
    data_exists = []
    has_data = 0
    # fix with None
    if(ddl_run):
      cur.execute("SELECT 1 FROM project;")
      if(cur.fetchone() is None):
        has_data = 0
      else:
        has_data = 1
      data_exists.append(has_data)
      cur.execute("SELECT 1 FROM grade;")
      if(cur.fetchone() is None):
        has_data = 0
      else:
        has_data = 1
      data_exists.append(has_data)
    else:
      data_exists.append(0)
      data_exists.append(0)
```

```{r check user uploads and clean}
UpdateProgress(renderVal, params$rendered_by_shiny)
#check to see if the user uploaded any docs
userData <- FALSE
availSurvBool <- FALSE
projSurvBool <- FALSE
eligibleBool <- FALSE

if(py$connected == TRUE){
  print("Connected to the DB this is where I am going to clean")
  if(py$ddl_run == 0 & !is.null(params$availSurv))
  {
    cleanAvailVec <- scheduling(params$availSurv)
    student <- cleanAvailVec[1]
    str(student)
    availSurvBool <- TRUE
  }
  if(!is.null(params$availSurv) | py$ddl_run == 1)
  {
    if(py$data_exists[1] == 0 & !is.null(params$projSurv) & !is.null(params$projInfo))
    {
      cleanProjVec <- projects(params$projSurv, params$projInfo) 
      project <- cleanProjVec[1]
      str(project)
      projSurvBool <- TRUE
    }
    if(py$data_exists[2] == 0 & !is.null(params$eligible))
    {
      cleanGradeVec <- grading(params$eligible)
      grade <- cleanGradeVec[1]
      str(grade)
      eligibleBool <- TRUE
    }
  }
}
UpdateProgress(renderVal, params$rendered_by_shiny)
```

```{python update or upload}
canQuery = False
if (connected == True and r.availSurvBool == True and ddl_run == False):
  print("Connected, user supplied data, and db is empty")
  # run ddl 
  # upload all...
  canQuery = True
elif (connected == True and ddl_run == True):
  if(r.projSurvBool == True):
    print("project data will be uploaded here")
  if(r.eligibleBool == True):
    print("grade data will be uploaded here")
  else:
    print("Connected, and database is setup. No data to upload or database already has data")
  canQuery = True
elif(connected == True and r.availSurvBool == False and ddl_run == False):
  print("ERROR: Connected but no data exists in the database, and no data provided to add")
else:
  print("Not connected")
```


# 20 Questions     
## Interst Level  
### Count the number of students interested in each project:  
_# of students with interest level greater than 5_  

### Count the number of times a project was selected as a top 3 choice:  

### Are there more students who marked low interest in all projects than students who marked high interest in all projects?  
_How to define low vs. high interest?_  

### Output a list of the 5 students who have the highest interest level for each project:  

### Create a scatterplot of interest level and confidence to complete each project:  

### What skills are students most interested in?  

## Availability  
### When are most students available?  
_Create a heatmap_

### Plot students time availability stratified by 1st choice of project:  

### Plot students time availability stratified by 2nd choice of project:  

### Plot students time availability stratified by 3rd choice of project:  

### What proportion of students perfer times before 10:00 Am?

### What proportion of students are available past 6:00 Pm? 

## Skills
### What are the 5 most common skills?  

### What are the 5 least common skills?  

### What is the most common skill among those who got straight A’s in their prerequisite courses?  
_Do they have to have an A in all pre-reqs? How many people actually did this?_  

### Return a list of the 5 students who are most confident in their ability to complete each project:  

### Create a histogram of the students confidence level in completing each project:  

### Create a barchart of the skills students are skilled in or familiar with:  

## Eligibility
### How many students do not meet eligibility criteria?   

### Where do skills not match with eligibility criteria?  
_May be hard to match skills and classes taken..._  

### What is the average confidence level of students who do not meet eligibility criteria for each project?  


```{python}
if(connected):
  conn.close()
```

